---
title: "Papa_WFP_oxygen_compare_to_discrete"
author: "Stace Beaulieu"
date: "2023-06-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(lubridate)
library(httr)
library(readxl)
library(dplyr)
library(ggplot2)
library(gridExtra)

```

## Compare profiler data with discrete data during turn around cruise

This R script plots discrete data from Alfresco against profiler data from Data Explorer ERDDAP during a turn-around cruise 


## Access Papa lower WFP data for oxygen during the transition between deployments

```{r}
# view in Data Explorer then select 2 days before and 2 days after in DE ERDDAP csvp 

# view in Data Explorer
# this is for between deployments 3 and 4
# https://dataexplorer.oceanobservatories.org/#metadata/104156/station/35/sensor/data?start=2016-06-22T00:00:00Z&end=2016-07-10T00:00:00Z&leg_clim_max=false&leg_clim_min=false
# this is for between deployments 4 and 5
# https://dataexplorer.oceanobservatories.org/#metadata/104156/station/35/sensor/data?start=2017-07-09T07:07:42Z&end=2017-07-27T02:59:49Z&leg_clim_max=false&leg_clim_min=false

# generate URL for csvp in ERDDAP for date range just including discrete data at HYPM
# this is for between deployments 3 and 4
# this is just the level 1 oxygen data in umol/L
# df <- read_csv("http://erddap.dataexplorer.oceanobservatories.org/erddap/tabledap/ooi-gp02hypm-wfp03-03-dostal000.csvp?time%2Cz%2Cmole_concentration_of_dissolved_molecular_oxygen_in_sea_water_profiler_depth_enabled&time%3E=2016-06-29T00%3A00%3A00Z&time%3C=2016-07-05T00%3A00%3A00Z&z%3E=-4067&z%3C=-2093") 

# this includes level 2 oxygen data in umol/kg
# note the Global Wire Following Profiler Salinity Drift Correction may be applicable to the Level 2 oxygen data (see OOI Discourse)

# this is for start deployment 1 26-28 July 2013
# ERDDAPurl <- "http://erddap.dataexplorer.oceanobservatories.org/erddap/tabledap/ooi-gp02hypm-wfp03-03-dostal000.csvp?time%2Cz%2Cmole_concentration_of_dissolved_molecular_oxygen_in_sea_water_profiler_depth_enabled%2Cmoles_of_oxygen_per_unit_mass_in_sea_water_profiler_depth_enabled%2Csea_water_practical_salinity_profiler_depth_enabled%2Csea_water_pressure_profiler_depth_enabled%2Csea_water_temperature_profiler_depth_enabled&time%3E=2013-07-26T00%3A00%3A00Z&time%3C=2013-07-29T00%3A00%3A00Z&z%3E=-4067&z%3C=-2093"

# this is for between deployments 1 and 2 13-23 June 2014 
# ERDDAPurl <- "http://erddap.dataexplorer.oceanobservatories.org/erddap/tabledap/ooi-gp02hypm-wfp03-03-dostal000.csvp?time%2Cz%2Cmole_concentration_of_dissolved_molecular_oxygen_in_sea_water_profiler_depth_enabled%2Cmoles_of_oxygen_per_unit_mass_in_sea_water_profiler_depth_enabled%2Csea_water_practical_salinity_profiler_depth_enabled%2Csea_water_pressure_profiler_depth_enabled%2Csea_water_temperature_profiler_depth_enabled&time%3E=2014-06-13T00%3A00%3A00Z&time%3C=2014-06-23T00%3A00%3A00Z&z%3E=-4067&z%3C=-2093"

# this is for between deployments 2 and 3 (i.e., for start 3) 5-6 June 2015
# ERDDAPurl <- "http://erddap.dataexplorer.oceanobservatories.org/erddap/tabledap/ooi-gp02hypm-wfp03-03-dostal000.csvp?time%2Cz%2Cmole_concentration_of_dissolved_molecular_oxygen_in_sea_water_profiler_depth_enabled%2Cmoles_of_oxygen_per_unit_mass_in_sea_water_profiler_depth_enabled%2Csea_water_practical_salinity_profiler_depth_enabled%2Csea_water_pressure_profiler_depth_enabled%2Csea_water_temperature_profiler_depth_enabled&time%3E=2015-06-04T00%3A00%3A00Z&time%3C=2015-06-07T00%3A00%3A00Z&z%3E=-4067&z%3C=-2093"

# this is for between deployments 3 and 4
ERDDAPurl <- "http://erddap.dataexplorer.oceanobservatories.org/erddap/tabledap/ooi-gp02hypm-wfp03-03-dostal000.csvp?time%2Cz%2Cmole_concentration_of_dissolved_molecular_oxygen_in_sea_water_profiler_depth_enabled%2Cmoles_of_oxygen_per_unit_mass_in_sea_water_profiler_depth_enabled%2Csea_water_practical_salinity_profiler_depth_enabled%2Csea_water_pressure_profiler_depth_enabled%2Csea_water_temperature_profiler_depth_enabled&time%3E=2016-06-29T00%3A00%3A00Z&time%3C=2016-07-05T00%3A00%3A00Z"

# this is for between deployments 4 and 5 
# ERDDAPurl <- "http://erddap.dataexplorer.oceanobservatories.org/erddap/tabledap/ooi-gp02hypm-wfp03-03-dostal000.csvp?time%2Cz%2Cmole_concentration_of_dissolved_molecular_oxygen_in_sea_water_profiler_depth_enabled%2Cmoles_of_oxygen_per_unit_mass_in_sea_water_profiler_depth_enabled%2Csea_water_practical_salinity_profiler_depth_enabled%2Csea_water_pressure_profiler_depth_enabled%2Csea_water_temperature_profiler_depth_enabled&time%3E=2017-07-14T00%3A00%3A00Z&time%3C=2017-07-18T00%3A00%3A00Z&z%3E=-4067&z%3C=-2093"

# this is for between deployments 6 and 7 25-27 Sep 2019 
# ERDDAPurl <- "http://erddap.dataexplorer.oceanobservatories.org/erddap/tabledap/ooi-gp02hypm-wfp03-03-dostal000.csvp?time%2Cz%2Cmole_concentration_of_dissolved_molecular_oxygen_in_sea_water_profiler_depth_enabled%2Cmoles_of_oxygen_per_unit_mass_in_sea_water_profiler_depth_enabled%2Csea_water_practical_salinity_profiler_depth_enabled%2Csea_water_pressure_profiler_depth_enabled%2Csea_water_temperature_profiler_depth_enabled&time%3E=2019-09-24T00%3A00%3A00Z&time%3C=2019-09-29T00%3A00%3A00Z"

df <- read_csv(ERDDAPurl)

# rename the columns to compare against discrete
df <- rename(df, profiler_oxygen_umol_L = `mole_concentration_of_dissolved_molecular_oxygen_in_sea_water_profiler_depth_enabled (micromol.L-1)`, profiler_oxygen_umol_kg = `moles_of_oxygen_per_unit_mass_in_sea_water_profiler_depth_enabled (micromol.kg-1)`)
# note the first few values on 2016-06-29 are zero; check if this is a fill value

# just need date for the plotting, better would be to create additional new column date
df$`time (UTC)` <- as_date(df$`time (UTC)`)
df <- rename(df, date = `time (UTC)`)

```


## Access OOI discrete data from Alfresco

```{r}

# this is for start deployment 1
# note the discrete sheet has no values in CTD Depth [m] thus these data need special treatment in order to plot as below
# url <- ('https://alfresco.oceanobservatories.org/alfresco/d/d/workspace/SpacesStore/aa1ad07d-ff3b-4aa1-bde2-e54f8504b333/Station_Papa-01_MV1309_Discrete_Summary.xlsx')

# this is for between deployments 1 and 2
# url <- ('https://alfresco.oceanobservatories.org/alfresco/d/d/workspace/SpacesStore/abc5c5b0-51f8-4758-afeb-0d11746f4b2c/Station_Papa-02_MV1404_Discrete_Summary.xlsx')

# this is for between deployments 2 and 3 (i.e., for start 3)
# url <- ('https://alfresco.oceanobservatories.org/alfresco/d/d/workspace/SpacesStore/5c61d4bf-73e0-4a79-8d47-2a1ecafe4769/Station_Papa-03_TN323_Discrete_Summary.xlsx')

# this is for between deployments 3 and 4
url <- ('https://alfresco.oceanobservatories.org/alfresco/d/d/workspace/SpacesStore/adf882e4-16cb-4d89-b89f-c96fcbc05e93/Station_Papa-04_RB1605_Discrete_Summary.xlsx')

# this is for between deployments 4 and 5
# url <- ('https://alfresco.oceanobservatories.org/alfresco/d/d/workspace/SpacesStore/e62d564f-c629-4f32-af03-6508b616452c/Station_Papa-05_SR1710_Discrete_Summary.xlsx')

# this is for between deployments 6 and 7
# url <- ('https://alfresco.oceanobservatories.org/alfresco/d/d/workspace/SpacesStore/31a9d169-abfe-4206-8e2c-781884ce1fa8/Station_Papa-07_SKQ201920S_Discrete_Summary.xlsx')

```

```{r}
# past experience if this chunk and the one above are not run separately, get Error
httr::GET(url, authenticate("guest", "guest"), write_disk(tf <- tempfile(fileext = ".xlsx")))
tf
discrete <- read_excel(tf, 1L) 

# need to remove -9999999 from `CTD Bottle Closure Time [UTC]` for the following workflow
discrete <- filter(discrete, `CTD Bottle Closure Time [UTC]` != -9999999)


# note that the following use of as.POSIXct() removes the time part but that all I need to retain is date
discrete$`CTD Bottle Closure Time [UTC]` <- as.POSIXct(discrete$`CTD Bottle Closure Time [UTC]`, tz="UTC")
discrete <- rename(discrete, date = `CTD Bottle Closure Time [UTC]`)
discrete$date <- as_date(discrete$date)

```

## Subset OOI discrete oxygen

```{r}
discreteoxygen <- discrete %>%
  select(date, "CTD Depth [m]", "CTD Oxygen [mL/L]", "Discrete Oxygen [mL/L]") %>%
  mutate(across("CTD Depth [m]":"Discrete Oxygen [mL/L]",as.numeric))

# for some reason I cannot filter
  # filter("CTD Depth [m]" > 2000) %>%
  # filter("Discrete Oxygen [mL/L]" > 0)
         
```

## Convert discrete oxygen for comparison to Data Explorer WFP oxygen

```{r}
discreteoxygen$`CTD Depth [m]` <- discreteoxygen$`CTD Depth [m]` * -1

# the following conversion is rounded from ICES (44.661)  and Bio-Argo (44.6596)
# Discrete oxygen as provided thru Data Explorer cruise data appears to use conversion 44.659599 from mL/L to umol/L (also rounds to 44.66); Data Explorer cruise data does not convert to umol/kg.
# Here is URL for Papa-04 Data Explorer cruise data as csv:
# https://servomatic9000.axiomalaska.com/geoserver/moving_platforms/ows?service=WFS&version=1.0.0&request=GetFeature&outputFormat=csv&typeName=moving_platforms:10bfff03-5ca6-5acc-a4c0-0ba8ee9c00f6
discreteoxygen <- mutate(discreteoxygen, discrete_oxygen_umol_L = `Discrete Oxygen [mL/L]` * 44.66)
discreteoxygen <- mutate(discreteoxygen, CTD_oxygen_umol_L = `CTD Oxygen [mL/L]` * 44.66)

# to convert to umol_kg the OOI BGC best practice is:
# Winkler (umol/kg) = Winkler (ml/l)*44.661*1000/water density
# the following uses constant 1025 kg/m3
# (conversion factor 43.5717 compares to NOAA WOD 43.570)
discreteoxygen <- mutate(discreteoxygen, discrete_oxygen_umol_kg = `Discrete Oxygen [mL/L]` * 44.661*1000/1025)
discreteoxygen <- mutate(discreteoxygen, CTD_oxygen_umol_kg = `CTD Oxygen [mL/L]` * 44.661*1000/1025)
# consider adjusting constant (e.g., to 1027.7 kg/m3 which is an approximate average density using an online calculator with Papa-04)
# better would be to use an equation to calculate density based on CTD bottle data temperature, salinity, pressure for bottles plotted here for discrete Winkler oxygen

# quick summary to check ranges
summary(discreteoxygen)

```

## Equate variables and overlay plots

```{r}
# rename profiler columns
profiler_to_plot <- df # initialize new data frame
profiler_to_plot <- profiler_to_plot %>%
  rename(depth = 'z (m)') %>%
  rename(oxygen_umol_L = profiler_oxygen_umol_L) %>%
  rename(oxygen_umol_kg = profiler_oxygen_umol_kg)

#rename discrete columns
discrete_to_plot <- discreteoxygen # initialize new data frame
discrete_to_plot <- discrete_to_plot %>%
  rename(depth = "CTD Depth [m]") %>%
  rename(oxygen_umol_L = discrete_oxygen_umol_L) %>%
  rename(oxygen_umol_kg = discrete_oxygen_umol_kg)

# ggplot(discrete_to_plot, aes(oxygen_umol_L, depth, color = date)) +
#    geom_point() +
#    xlim(50, 175) +
#    ylim(-4300, -2000)
# 
# ggplot(profiler_to_plot, aes(oxygen_umol_L, depth, color = date)) +
#    geom_point() +
#    xlim(50, 175) +
#    ylim(-4300, -2000)

# superimpose plots Level 1 oxygen umol_L
p1 <- ggplot(discrete_to_plot, aes(oxygen_umol_L, depth, color = date)) +
   geom_point() +
   geom_point(data = profiler_to_plot) +
   xlim(30, 175) +
   ylim(-4250, -2000)

# superimpose plots Level 2 oxygen umol_kg
# note I have not yet applied the Global Wire Following Profiler Salinity Drift Correction (see OOI Discourse)
# broader x axes limits than 2018 SME ipynb
p2 <- ggplot(discrete_to_plot, aes(oxygen_umol_kg, depth, color = date)) +
   geom_point() +
   geom_point(data = profiler_to_plot) +
   xlim(30, 150) +
   ylim(-4250, -2000)

# two-panel plot side-by-side
grid.arrange(p1, p2, nrow = 1)

```

## Test plots using Level 1 oxygen umol_L

```{r}

# plot CTD bottle data
ggplot(discreteoxygen, aes(CTD_oxygen_umol_L,`CTD Depth [m]`, color = date)) +
  geom_point() +
  xlim(30, 175) +
  ylim(-4250, -2000)


```

